---
title: "Key Driver Analysis"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Key Driver Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(YouAnalyser)
library(haven)
```

# Recommended Workflow

The following minimal and basic workflow is recommended for performing Key Driver Analysis (KDA) using the `YouAnalyser` package. This workflow includes the following steps:

1. Exploratory Data Analysis (EDA)
2. Key Driver Analysis (KDA)
3. Visualize and interpret the results

It is assumed that the data is already in a `haven::labelled()` format. If your data is not in this format, you may need to perform some data processing steps to convert it before using `YouAnalyser`. For an example of how to do this, please refer to the `vignette("dp", package = "YouAnalyser")` vignette on Data Processing.

```{r basic-workflow-example, eval = FALSE}
#---- 1. EDA -------------------------------------------------------------------

# Load your data
# myData <- haven::read_sav("path/to/your/data.sav")
# Here, we will use the example data included in the package:
myData <- YouAnalyser::bkw_processed

# Get an overview of the data. Confirm there are no unexpected values
# (e.g., values like "99") and no missing values.
eda_summary(myData)

# Inspect correlations among the variables.
# Are all predictors positively correlated with the outcome variable?
# Are there any high correlations among the predictors that may indicate
# multicollinearity issues?
eda_corrs <- eda_correlation(myData)
print(eda_corrs$p)

#---- 2. KDA -------------------------------------------------------------------

# Perform Key Driver Analysis
kda <- kda_regression(
  data = myData,
  outcome = "F600",
  predictors = paste0("F800_", 1:8)
)

#---- 3. Visualize and interpret results ---------------------------------------

# Inspect the Importance Performance Matrix Analysis (IPMA) plot
print(kda$plots$ipma_scatterPlot$p)

# Save the IPMA plot as a file
ya_save_plot(
  plot = kda$plots$ipma_scatterPlot$p,
  file_path = ya_choose_file_path("myPlotName.jpeg"),
  width = 30,
  height = 20
)

# If you need a PowerPoint report:

# 1. Export the KDA results into a formatted Excel template:
kda_save_data_for_chart(
  ipma_scatterPlot_data = kda$plots$ipma_scatterPlot$d,
  file_path = ya_choose_file_path("myKDAresults.xlsx")
)

# 2. Copy the PowerPoint template provided within YouAnalyser to a
# location of your choice:
kda_copy_pptx_template(
  file_path = ya_choose_file_path("myKDAReport.pptx")
)
```

For the remainder of this vignette, we will focus on the different options for conducting the KDA step (step 2 in the workflow above). For a more detailed walkthrough of the EDA step, please refer to the `vignette("eda", package = "YouAnalyser")` vignette on Exploratory Data Analysis.

# The `kda_regression()` Function

The `kda_regression()` function is the main work horse for conducting KDA. It is a wrapper function for all the necessary steps (subfunctions) to conduct a KDA, including fitting a regression model, calculating variable importance and performance measures, and creating the relevant plots. The function has several arguments that allow you to customize the analysis, such as the choice of regression model, variable importance method, and diagnostic options. You can get help on the function and its arguments by running `help("kda_regression")` or `?kda_regression` in your R console.

You need to provide at least the following arguments to the `kda_regression()` function: 

- Option 1: `data`, `outcome`, and `predictors`
  - `data`: A data frame containing the outcome and predictors. The data needs to be in a `haven::labelled()` format, as is typically the case for SPSS data. If it is not, refer to the `vignette("dp", package = "YouAnalyser")` vignette on Data Processing for how to convert your data into a labelled format.
  - `outcome`: A single string naming the outcome variable.
  - `predictors`: A character vector of predictor variable names. You can use e.g. `paste0("F800_", 1:8)` to specify a sequence of variables with similar names.
- Option 2: `model`: Instead of providing the `data`, `outcome`, and `predictors` arguments, you can also directly provide a fitted regression model object from `lm()` or `glm()` to the `model` argument (intended for advanced users.)

The `kda_regression()` function also has several other options. These have recommended default values and do not need to be changed in most cases:

- `diagnostics = FALSE`: A logical indicating whether to compute model diagnostics. This is not necessary for the KDA itself, but can be useful for checking the assumptions of the regression model. If set to `TRUE`, diagnostic plots will be included in the output.
- `importance_method = "auto"`: The method to calculate variable importance. The default "auto" will conduct a dominance analysis if the number of predictors is less than 15. Otherwise, it will use the "jrw" method, which is a computationally efficient method for calculating variable importance in larger models. See below for more information on the available variable importance methods.
- `importance_barPlots_args = list()`: A list of additional arguments passed to the importance bar plot function. See `kda_importance_barPlot()` for details.
- `performance_barPlot_args = list()`: A list of additional arguments passed to the performance bar plot function. See `kda_performance_barPlot()` for details.
- `ipma_scatterPlot_args = list()`: A list of additional arguments passed to the IPMA scatter plot function. See `kda_ipma_scatterPlot()` for details.

Regarding the `_args` arguments, you can use these to customize the appearance of the plots. For example, you can change the default color of the performance_barPlot from YouGov Red 1 to Teal 1 like this: `performance_barPlot_args = list(color = yougov_colors[["Teal 1"]])`. For more information on the available arguments, please refer to the documentation of the respective plotting functions.

