---
title: "Key Driver Analysis: Article"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(YouAnalyser)
library(haven)
```


# Exploratory Data Analysis (EDA)

It is good practice to start with an exploratory data analysis (EDA) to understand the structure of the data, check for missing values, and identify any potential issues before performing Key Driver Analysis (KDA). `YouAnalyser` provides conventient functions for EDA, all starting with *eda_*: `eda_summary()` and `eda_correlations()`. These functions can help you get a quick overview of your data and identify any necessary preprocessing steps. But first, we need some data!

## Reading in the data

Note that the data should be in [haven::labelled()] format for `YouAnalyser` to work properly, which is common for survey data. If your data is in a different format, you may need to convert it before using `YouAnalyser`.

We start by loading the data. The `YouAnalyser` package includes a sample dataset called `bwk.sav`, which is in SPSS format. You can read in your own dataset using the `haven::read_sav()` function, e.g.:

```{r eval = FALSE}
myOwnData <- haven::read_sav("path/to/your/ownFile.sav")
```

We can read in the included sample dataset as follows:

```{r eval = FALSE}
example_data <- haven::read_sav(system.file(
  "extdata",
  "bkw.sav",
  package = "YouAnalyser"
))

# This will be equivalent to `example_data <- YouAnalyser::bkw_missings` but using `haven::read_sav()` allows us to see how to read in our own data in the same format.
```

```{r include = FALSE}
example_data <- YouAnalyser::bkw_missings
```

`example_data` contains our outcome variable `F600` and 14 predictor variables `F800_1` to `F800_14`. To get more information about this included dataset, you can run `?YouAnalyser::bkw_missings` in the R console:

```{r eval = FALSE}
help(YouAnalyser::bkw_missings)
```

We will select only the variables of interest:

```{r}
variables_of_interest <- c("F600", paste0("F800_", 1:14))
example_data <- example_data |>
  dplyr::select(all_of(variables_of_interest))
```

## Get a summary of the data

The `eda_summary()` function provides a summary of the data, including the number of observations, number of missing values, and basic statistics for each variable. This can help you understand the distribution of your data and identify any potential issues.

```{r eval = FALSE}
eda_summary(example_data)
```

This will open a new browser window with the summary of the data that will look something like this:

```{r echo = FALSE, results = 'asis'}
summarytools::dfSummary(
  example_data,
  plain.ascii = FALSE,
  style = "grid",
  graph.magnif = 0.82,
  varnumbers = FALSE,
  valid.col = FALSE,
  tmp.img.dir = "/tmp"
) |>
  print(method = "render", headings = FALSE)
```

We see that there are 4307 missing values in the outcome and predictor variables. Lets remove these missing values and save the new dataset as `example_data_clean`:

```{r}
example_data_clean <- example_data |>
  tidyr::drop_na()
```

## Check correlations

We can also have a look at the correlation matrix of the variables using the `eda_correlations()` function. This can help you identify any potential multicollinearity issues among the independent variables as well as the strength and directionof the relationships between the independent variables and the dependent variable. High correlations among independent variables can indicate multicollinearity, which can affect the stability and interpretability of the KDA results. Moreover, we would like to see that the independent variables are correlated **positivly** with the dependent variable, as KDA is based on the idea of identifying key drivers that have a positive impact on the outcome.

```{r}
corrs <- eda_correlation(example_data_clean, correlation_type = "pearson")
```

A correlation matrix plot is saved in `corrs$p`, which will look like this:

```{r fig.width=12, fig.height=12, out.width="100%"}
corrs$p
```

We see that there are some high correlations among the independent variables, which may indicate multicollinearity. However, we also see that all independent variables are positively correlated with the dependent variable `F600`, which is a good sign for performing KDA.

# Key Driver Analysis (KDA)

Now the real fun starts! [YouAnalyser::kda_regression()] is the main function for performing Key Driver Analysis. There are two ways to use this function: (1) by providing the raw data and specifying the outcome variable and predictor variables, or (2) by providing a fitted model object. The first approach is more straightforward and is recommended for most users, while the second approach allows more advanced users to use some shortcuts, as we will show below.

The recommended approach in its minimal form looks like this:

```{r eval = FALSE}
kda <- kda_regression(
  data = example_data_clean,
  outcome = "F600",
  predictors = paste0("F800_", 1:14)
)
```

Alternatively, we can first fit a model and then provide the fitted model object to [YouAnalyser::kda_regression()]. This for instance allows for short cuts like `outcome ~ .` in the model formula. If you do not recognize this syntax, it is perfectly fine to ignore it and use the first approach as shown above. The second approach would look like this:

```{r eval = FALSE}
myModel <- lm(F600 ~ ., data = example_data_clean)
kda <- kda_regression(model = myModel)
```

Note that it is only possible to use models fitted with `lm()` and `glm()` without interaction terms! (Interactions make it very difficult to calculate usefull importance measures for KDA).

```{r include = FALSE}
# To save time, we can save the results of the KDA regression to a file and read it in again when needed, instead of running the regression every time we knit the vignette. This is especially useful if the regression takes a long time to run, as is the case for the dominance analysis method with 14 predictors.
if (FALSE) {
  kda <- kda_regression(
    data = example_data_clean,
    outcome = "F600",
    predictors = paste0("F800_", 1:14)
  )
  saveRDS(kda, file = "./kda_dominanceAnalysis.rds")
} else {
  kda <- readRDS("../kda_dominanceAnalysis.rds")
}
```

It is highly recommended to have a look at the help documentation of [YouAnalyser::kda_regression()] to see all the different options for performing KDA, including different methods for calculating importance measures and different ways to visualize the results as well as all the available output. You can access the help documentation by running `?YouAnalyser::kda_regression` in the R console.

Let's have a look at the different outputs generated by [YouAnalyser::kda_regression()]: